Laurentia in the earliest Neoproterozoic: new paleomagnetic constraint from the Jacobsville Formation

### 2022-07-23 Manuscript/Manuscript.tex
Updates from Overleaf

\begin{abstract}
Coeval with the cessation of the ca. 1.1 Ga intracontinental Midcontinent Rift was the onset of the ca. 1090-980 Ma Grenvillian Orogeny along Laurentia's eastern margin. Far-field compressional stresses of the orogeny propagated into the interior of Laurentia and inverted the Midcontinent Rift, deforming the post-rift Jacobsville Formation whose maximum depositional age has been constrained to be ca. 990 Ma. We develop a new paleomagnetic pole from red beds of the Jacobsville Formation in northern Michigan, USA. High-resolution thermal demagnetization experiments resolve detrital remanent magnetizations that pass a conglomerate test. An inclination corrected paleomagnetic mean pole position from the Jacobsville Formation confirm continued plate motion of Laurentia across the equator in the late Mesoproterozoic to early Neoproterozoic. The new pole from the interior of Laurentia indicates the paleomagnetic poles developed from the Grenville Province most likely record coherent plate motion after major continental collision during the initial assembly of the supercontinent Rodinia.
\end{abstract}

>>>>

\begin{abstract}
The paleogeography of Laurentia in the late Proterozoic is central to global paleogeography due to its central position in the supercontinent Rodinia. We develop a new paleomagnetic pole from red beds of the ca. 990 Ma Jacobsville Formation which serves as a crucial anchor point for Laurentia's position in the earliest Neoproterozoic. High-resolution thermal demagnetization experiments resolve detrital remanent magnetizations that pass a conglomerate test. An inclination corrected paleomagnetic mean pole position from the Jacobsville Formation confirm continued plate motion of Laurentia across the equator in the late Mesoproterozoic to early Neoproterozoic. The new pole from the interior of Laurentia indicates the plate slowed down significantly following the onset of the Grenvillian orogenesis on its eastern margin. This pole's position and age indicate the paleomagnetic poles developed from the Grenville Province most likely record coherent plate motion after major continental collision during the initial assembly of the supercontinent Rodinia.
\end{abstract}
